
```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(tidyr)
library(foreach)
library(som.nn)
library(mcomsimr)
```



Function definition calculate the population covariance ##


```{r function pop covariance}



# Function to calculate the population covariance (i.e. not an unbiased
# sample covariance with denominator n-1, which is the default in R).
covp <- function(x,y){
	n <- length(x)
	return( (n-1) * cov(x,y) / n )
}


```


Function definition for Beverton-Holt competition growth model##


```{r function BV holt growth}


##N = pop size##
## a = competition coefficient####

# Function definition for Beverton-Holt competition
bevHoltC <- function(N,a) {
	return( 1 / ( 1 + a*(sum(N))))
}

###represent per capita pop growth w competition from self and other spp### 

##need to multiply this by R (intrinsic pop growth rate) to get Nt+1###

```



Function definition for deterministic pop growth##


```{r function population growth}


## a = competition coefficients####
## R = intrinsic population growth rate###



b <- function(R,current, patches, species, a) {
	birth <- matrix(NA, nrow=species, ncol=patches)
	for (p in 1:patches) {
		birth[,p] <- R[,p]*current[,p]*bevHoltC(current[,p],a)
	}
	return(birth)
}

```



Function definition deterministic model run BV holt pop growth and random movement btwn patches #

```{r model function definition random movement (Shoemaker 2016)}

# Function definition for deterministic model run ### random movement between patches##


rdet <- function(R,d,a,patches,species,data,time) {
	
	for (t in 1:(time-1)){
		
		# Number of individuals in each patch after pop growth but before dispersal###
		birth <- b(R, data[,,t],patches, species, a)
		
		# dispersal
		migrate <- d*birth
		stay <- birth - migrate
		
		totalmigrate <- rep (NA, species)
		for (s in 1:species) {
			totalmigrate[s] <- sum(migrate[s,])
		}
		
		# determine where each individual migrates
		migratelocation <- matrix(NA, nrow=species, ncol=patches)
		for (s in 1:species) {
			migratelocation[s,] <- rep(totalmigrate[s]/patches, patches)
		}
		
		# Number of individuals in each patch after individuals have moved
		data[,,t+1] <- stay + migratelocation	
		
	}
return(data)
}






```


Function definition for deterministic model run for determining Chessonian coexistence

```{r model function definition Chessonian coexistence (random movement)}

# Function definition for deterministic model run for determining Chessonian coexistence 
# (neutral, species sorting, or mass effects)
rdetChesson <- function(R,d,a,patches,species,data,invader,invadeTotal,time) {
	
	for (t in 1:(time-1)){
		
		# birth
		
		birth <- b(R, data[,,t],patches, species, a)
		
		# dispersal
		migrate <- d*birth
		stay <- birth - migrate
		
		totalmigrate <- rep (NA, species)
		for (s in 1:species) {
			totalmigrate[s] <- sum(migrate[s,])
		}
		
		# determine where each individual migrates
		migratelocation <- matrix(NA, nrow=species, ncol=patches)
		for (s in 1:species) {
			migratelocation[s,] <- rep(totalmigrate[s]/patches, patches)
		}
		
		# Determine number of individuals in each patch
		data[,,t+1] <- stay + migratelocation	

		
		# Reset the invader
		data[invader,,t+1] <- data[invader,,t+1] / sum(data[invader,,t+1]) * invadeTotal
		
	}
return(data)
}


```


Function definition deterministic model run fitness-directed dispersal (disp is a function of R) #

```{r model function definition fitness-directed movement global}

# Function definition for deterministic model run ### fitness-dependent movement between patches##

rdetdirected <- function(R,d,a,pref,patches,species,data,time) {
	
	for (t in 1:(time-1)){
		
		# Number of individuals in each patch before dispersal###
		birth <- b(R, data[,,t],patches, species, a)
		
		# dispersal
		migrate <- d*birth
		stay <- birth - migrate
		
		totalmigrate <- rep (NA, species)
		for (s in 1:species) {
			totalmigrate[s] <- sum(migrate[s,])
		}
		
		# probability of dispersing to that patch depends on R ###
		probpatch <- matrix(NA, nrow=species, ncol=patches)
		for (s in 1:species) {
			 for(p in 1:patches){
		  probpatch[s,p] <- exp(pref*R[s,p])  ### probability of going to patch exponential to R (pref = preference of that patch (positive constant))####
			 }
		}
		
		# determine where each individual migrates
		migratelocation <- matrix(NA, nrow=species, ncol=patches)
		for (s in 1:species) {
			 for(p in 1:patches){
		  migratelocation[s,p] <- (probpatch[s,p]*totalmigrate[s])/sum(probpatch[s,])
			 }
		}
		
		
		# Number of individuals in each patch after individuals have moved
		data[,,t+1] <- stay + migratelocation	
		
	}
return(data)
}






```




```{r model function definition fitness-directed movement global tweaked version}

# Function definition for deterministic model run ### fitness-dependent movement between patches##

rdetdirected <- function(R,d,a,pref,patches,species,data,time) {
	
	for (t in 1:(time-1)){
		
		# Number of individuals in each patch before dispersal###
		birth <- b(R, data[,,t],patches, species, a)
		
		# dispersal
		migrate <- d*birth
		stay <- birth - migrate
		
		totalmigrate <- rep (NA, species)
		for (s in 1:species) {
			totalmigrate[s] <- sum(migrate[s,])
		}
		
		# probability of dispersing to that patch depends on R ###
		probpatch <- matrix(NA, nrow=species, ncol=patches)
		for (s in 1:species) {
			 for(p in 1:patches){
		  probpatch[s,p] <- exp(pref*R[s,p])/sum(exp(pref*R[s,]))  ### probability of going to patch exponential to R (pref = preference of that patch (positive constant))####
			 }
		}
		
		# determine where each individual migrates
		migratelocation <- matrix(NA, nrow=species, ncol=patches)
		for (s in 1:species) {
			 for(p in 1:patches){
		  migratelocation[s,p] <- (probpatch[s,p]*totalmigrate[s])
			 }
		}
		
		
		# Number of individuals in each patch after individuals have moved
		data[,,t+1] <- stay + migratelocation	
		
	}
return(data)
}






```




Function definition for deterministic model run for determining Chessonian coexistence: fitness-directed dispersal

###fix this to incorporate pref from above### 

```{r model function definition Chessonian coexistence fitness-directed movement}

# Function definition for deterministic model run for determining Chessonian coexistence 
# (neutral, species sorting, or mass effects)
rdetdirectedChesson <- function(R,d,a,patches,species,data,invader,invadeTotal,time) {
	
	for (t in 1:(time-1)){
		
		# birth
		
		birth <- b(R, data[,,t],patches, species, a)
		
		# dispersal
		migrate <- d*birth
		stay <- birth - migrate
		
		totalmigrate <- rep (NA, species)
		for (s in 1:species) {
			totalmigrate[s] <- sum(migrate[s,])
		}
		
		
		# probability of dispersing to that patch depends on R ###
		probpatch <- matrix(NA, nrow=species, ncol=patches)
		for (s in 1:species) {
			 for(p in 1:patches){
		  probpatch[s,p] <- exp(5*R[s,p])  ### probability of going to patch exponential to R (5 is arbitrary)####
			 }
		}
		
		# determine where each individual migrates
		migratelocation <- matrix(NA, nrow=species, ncol=patches)
		for (s in 1:species) {
			 for(p in 1:patches){
		  migratelocation[s,p] <- (probpatch[s,p]*totalmigrate[s])/sum(probpatch[s,])
			 }
		}
		
		# Determine number of individuals in each patch
		data[,,t+1] <- stay + migratelocation	

		
		# Reset the invader
		data[invader,,t+1] <- data[invader,,t+1] / sum(data[invader,,t+1]) * invadeTotal
		
	}
return(data)
}


```



Function definition deterministic model # immigrants depends on density of patch (dd -) #


```{r model function definition density-dependent (-) immigration, prob D}

# Function definition for deterministic model run ### negative density-dependent immigration between patches##


rdetddi <- function(R,d,a,patches,species,data,time) {
	
	for (t in 1:(time-1)){
		
		# Number of individuals in each patch after pop growth but before dispersal###
		birth <- b(R, data[,,t],patches, species, a)
   	birth[is.na(birth)] <- 0
		
		# dispersal
		migrate <- d*birth
		stay <- birth - migrate
		
		##total number of each spp migrating###
		totalmigrate <- rep (NA, species)
		for (s in 1:species) {
			totalmigrate[s] <- sum(migrate[s,])
		}
		

		# probability of dispersing to that patch depends on density of individuals staying ###
		probpatch <- matrix(NA, nrow=species, ncol=patches)
		for (s in 1:species) {
			 for(p in 1:patches){
		  probpatch[s,p] <- exp(-1*5*stay[s,p]/sum(stay[s,]))  ### neg exp of density #
			 }
		}
		
		# determine where each individual migrates
		migratelocation <- matrix(NA, nrow=species, ncol=patches)
		for (s in 1:species) {
			 for(p in 1:patches){
		  migratelocation[s,p] <- (probpatch[s,p]*totalmigrate[s])/sum(probpatch[s,])
			 }
		}
		
		
		
		# Number of individuals in each patch after individuals have moved
		data[,,t+1] <- stay + migratelocation	
		data[,,t+1][is.na(data[,,t+1])] <- 0
		
	}
return(data)
}





```

Function definition deterministic model run for determining Chessonian coexistence:# immigrants depends on density of patch (dd - ) #


```{r model function definition Chessonian coexistence density-dependent (-) immigration}

# Function definition for deterministic model run for determining Chessonian coexistence 
# (neutral, species sorting, or mass effects)
rdetddiChesson <- function(R,d,a,patches,species,data,invader,invadeTotal,time) {
	
	for (t in 1:(time-1)){
		
		# birth
		
		birth <- b(R, data[,,t],patches, species, a)
		birth[is.na(birth)] <- 0
		
		# dispersal
		migrate <- d*birth
		stay <- birth - migrate
		
		totalmigrate <- rep (NA, species)
		for (s in 1:species) {
			totalmigrate[s] <- sum(migrate[s,])
		}
		
		
			# probability of dispersing to that patch depends on density of individuals staying ###
		probpatch <- matrix(NA, nrow=species, ncol=patches)
		for (s in 1:species) {
			 for(p in 1:patches){
		  probpatch[s,p] <- exp(-1*5*stay[s,p]/sum(stay[s,]))  ### neg exp of density ####
			 }
		}
		
		# determine where each individual migrates
		migratelocation <- matrix(NA, nrow=species, ncol=patches)
		for (s in 1:species) {
			 for(p in 1:patches){
		  migratelocation[s,p] <- (probpatch[s,p]*totalmigrate[s])/sum(probpatch[s,])
			 }
		}
		
		
		# Number of individuals in each patch after individuals have moved
		data[,,t+1] <- stay + migratelocation	
    data[,,t+1][is.na(data[,,t+1])] <- 0
		
		# Reset the invader
		data[invader,,t+1] <- data[invader,,t+1] / sum(data[invader,,t+1]) * invadeTotal
		
	}
return(data)
}


```


Function definition deterministic model # immigrants depends on density of patch (dd +) #


```{r model function definition density-dependent (+) immigration, prob D}

# Function definition for deterministic model run ### negative density-dependent immigration between patches##


rdetddi_pos <- function(R,d,a,patches,species,data,time) {
	
	for (t in 1:(time-1)){
		
		# Number of individuals in each patch after pop growth but before dispersal###
		birth <- b(R, data[,,t],patches, species, a)
   	birth[is.na(birth)] <- 0
		
		# dispersal
		migrate <- d*birth
		stay <- birth - migrate
		
		##total number of each spp migrating###
		totalmigrate <- rep (NA, species)
		for (s in 1:species) {
			totalmigrate[s] <- sum(migrate[s,])
		}
		

		# probability of dispersing to that patch depends on density of individuals staying ###
		probpatch <- matrix(NA, nrow=species, ncol=patches)
		for (s in 1:species) {
			 for(p in 1:patches){
		  probpatch[s,p] <- exp(5*stay[s,p]/sum(stay[s,]))  ### positive exp of density #
			 }
		}
		
		# determine where each individual migrates
		migratelocation <- matrix(NA, nrow=species, ncol=patches)
		for (s in 1:species) {
			 for(p in 1:patches){
		  migratelocation[s,p] <- (probpatch[s,p]*totalmigrate[s])/sum(probpatch[s,])
			 }
		}
		
		
		
		# Number of individuals in each patch after individuals have moved
		data[,,t+1] <- stay + migratelocation	
		data[,,t+1][is.na(data[,,t+1])] <- 0
		
	}
return(data)
}





```



```{r model function definition Chessonian coexistence density-dependent (+) immigration}

# Function definition for deterministic model run for determining Chessonian coexistence 
# (neutral, species sorting, or mass effects)
rdetddiChesson_pos <- function(R,d,a,patches,species,data,invader,invadeTotal,time) {
	
	for (t in 1:(time-1)){
		
		# birth
		
		birth <- b(R, data[,,t],patches, species, a)
		birth[is.na(birth)] <- 0
		
		# dispersal
		migrate <- d*birth
		stay <- birth - migrate
		
		totalmigrate <- rep (NA, species)
		for (s in 1:species) {
			totalmigrate[s] <- sum(migrate[s,])
		}
		
		
			# probability of dispersing to that patch depends on density of individuals staying ###
		probpatch <- matrix(NA, nrow=species, ncol=patches)
		for (s in 1:species) {
			 for(p in 1:patches){
		  probpatch[s,p] <- exp(5*stay[s,p]/sum(stay[s,]))  ### postivie exp of density ####
			 }
		}
		
		# determine where each individual migrates
		migratelocation <- matrix(NA, nrow=species, ncol=patches)
		for (s in 1:species) {
			 for(p in 1:patches){
		  migratelocation[s,p] <- (probpatch[s,p]*totalmigrate[s])/sum(probpatch[s,])
			 }
		}
		
		
		# Number of individuals in each patch after individuals have moved
		data[,,t+1] <- stay + migratelocation	
    data[,,t+1][is.na(data[,,t+1])] <- 0
		
		# Reset the invader
		data[invader,,t+1] <- data[invader,,t+1] / sum(data[invader,,t+1]) * invadeTotal
		
	}
return(data)
}


```



```{r plot results random movement , echo=FALSE}

###turn array into dataframe###

dispdatarandom <- as.data.frame.table(dispscenariosrandom122)

colnames(dispdatarandom) <- c("species", "patch", "timestep", "d", "N")

dispdatarandom$species<-as.numeric(dispdatarandom$species)
dispdatarandom$patch<-as.numeric(dispdatarandom$patch)
dispdatarandom$timestep<-as.numeric(dispdatarandom$timestep)
dispdatarandom$d <- with(dispdatarandom, gsub("d=","", d))


##assign R values to patches###
 
species1<-as.numeric(dispdatarandom$species)
patch1<-as.numeric(dispdatarandom$patch)

Rvals <- vector("numeric", length=length(species1))
for (i in 1:length(species1)){
   s<-species1[i]
   p<-patch1[i]
   Rvals[i]<-c(R[s,p])
}


##merge back to dataframe###

dispdatarandom2<-cbind(dispdatarandom, Rvals)

##column for patch type###

dispdatarandom3<-dispdatarandom2 %>% mutate(patchtype=ifelse(species==1 & Rvals>mean(dispdatarandom2$Rvals),"Source sp 1 / Sink sp 2", ifelse(species==2 & Rvals < mean(dispdatarandom2$Rvals),"Source sp 1 / Sink sp 2", "Source sp 2 / Sink sp 1")))

###sum across patch types####

sumpatchrandom<-dispdatarandom3 %>%
  group_by(species, patch, timestep, d, patchtype) %>%
  summarise(N = sum(N))

sumpatchrandom$species=as.factor(sumpatchrandom$species)
sumpatchrandom$timestep<-as.numeric(sumpatchrandom$timestep)

#plot##
randommovement2spp<-ggplot(sumpatchrandom ,aes(x=timestep,y=N,colour=species, shape=patchtype), show.legend = F) +geom_line(aes(linetype=patchtype), size=1.5)+facet_grid(d~species, labeller = label_both)+xlab("Time")+ylab("Abundance")+ggtitle("Passive dispersal btwn patches")


###
```


